\TITLE

### 2021-01-12 response-v3-content.tex

\bigskip

\parindent 0pt We are happy to resubmit our paper and are very thankful for the great reviews we received.

If possible we would like to acknowledge the reviewers by name in the
acknowledgement section, but this depends on them and you.

We include the following comments

\section*{Additional actions taken}

@@

\section*{Reviewer 1}

The revised manuscript is much smoother than the first version, but I
still feel that less than sufficient care was put into the
writing. The reviewer pointed some in the pdf report. I have some
additional editorial suggestions:

\bigskip

  \REVIEW{DONE}{Submit the point-by-point response to the
    reviewer's comments and my comments in a pdf file.}{The original
    submisison web page had a field to enter the comments as text
    which we used. We just added ``see attached PDF'' to make it
    easier for reviewers as suggested by the editor.}

@@

  \REVIEW{DONE} {Needs page number (line number) for ease of
    referencing in the review.}
  {Page numbers and line numbers were
    activated with the review style from the new JDS \LaTeX{}
    template}

  \REVIEW{DONE} {Do not italicize in the text.}
  {All italicized text
    have been removed.  Formulas enclosed in dollar signs are in
    italicized text as is default for the JDS document style.}

@@

  \REVIEW{DONE} {AICov needs to be defined in the abstract.}
  {We added ``We present the architecture of an artificial
    intelligence enhanced COVID-19 analysis (in short AICov)''
  in the abstract. We also defined COVID-10 in the abstract.}

\REVIEW{DONE}{Acronyms need to be defined at their first occurrences
  (e.g., AI in the intro). A table of acronyms is included in table 1,
  we added the following acronyms to the paper}
  {All acronyms have been defined at their first usage. In addition
    we have added a section Acronyms in which all acronyms are listed
    to make it easier for the reader to look them up in case they are
    used later on in the paper. All risk factors have been capitalized.
    The duplicated Risk factors were removed.
  }

\REVIEW{DONE}{Figure 2: please present the equations as equations. The
  notations used in the equation need a full check (e.g., h is used as
  both the dimension and vector). I'd like to see a clear distinction
  between inputs/outputs and parameters. I'd suggest that the inputs
  be set up before the equations are presented.}
  { We have replaced the h with a capital H, We have started our explanation of the equations with the input and output vectors, we have removed the equations from the image and displayed them as separate equations using the LaTeX equation environment with numbers. }

@@

\REVIEW{DONE}{Regenerate the figures into pdf format (the current
  jpg/png figures blur when zoomed in).}
  {All images have been included as PDF images}

\REVIEW{DONE}{Table 2: consider reformatting it; it would not fit the
  page width in print.}
  { Table 2 has been reformatted. It was previously fit in the column
    with resize box. While rotating the headers the font is
    increased. The table still fits in the column as it still uses
    resizebox. Thus it will fit in the print.}

@@

\REVIEW{DONE}{Table 3: tidy the headers.}
  { The headers have been cleaned up and linebreaks were introduced so
    the resizebox of the table presents the table now larger.}

\REVIEW{DONE}{Figure 5-6: can be combined to one figure with two
  panels (using ggplot2 and save the space of the labels of the
  horizontal axis).  }
  {The figures were combined as subfigures into Figure \ref{fig:magic-1} (a) and (b)}

\REVIEW{DONE}{Figure 7-8: can be one figure with two panels; pop
  density 2010 needs to be cleaned.}
    {The figures were combined as subfigures into Figure \ref{fig:error-death-popdensity} (a) and (b)}

\REVIEW{DONE}{Figure 9-14: combine into one figure.}
  {

% \todo{the lines are missing in the images showing the NONE line}

@@

}

\REVIEW{DONE}{ Page 7, Table 1, is “CHD” also a percent? If yes
  please modify the description; Percent- blacks and black percent
  both correspond to “Percent of blacks in the population”. Both seem
  to have been used according to Table 3, with difference in the four
  model performance measurements. Do the authors have an explanation
  for this?  }
%
{ (a) The CHD and other values in the table have been updated wherever
  they are given in percent.

  (b) There were two different information sources that we originally
  used indicated by the capitalization. However since the submission
  we have improved this data and have now merged all data and
  corrected information for these duplicates, while only presenting
  the one that is more accurate.

  (c) We have fixed all values. Furthermore we have split the table with the values in two tables fro deaths and for cases to relate them more easily for the reader to the box diagrams. This makes the display of the data consistent and the tables and figures match. 
}

@@

\REVIEW{DONE}{Page 8, line 5 of Section 5.1, the abbreviation LSTM has
  already appeared before, and the full name can be removed.}
{All acronyms have been defined at first usage. A table has been added
  at the end to make it easier for the reader to look them up.}

\REVIEW{DONE}{What do the two panels in Figure 4 represent? Please give
  clearer description either in the text or figure caption.
}
  { Better descriptions have been added. }

\REVIEW{DONE}{Section 5.2, paragraph 2 line 6, no need to capitalize for
  the word ``Samples''. Line 5 from the bottom, no need to capitalize
  for the word “Insurance”. Please be careful with your writing.
}
{The capitalization has been corrected. The paper has been reviewed by
  2 native English speakers. In addition 2 grammar checking programs
  were used on the entire paper. One of the reviewers serves also as
  editor in other journals.}

\REVIEW{DONE}{Section 5.2, the ``order for cases sorted by minimum model
  order'' and ``order for deaths sorted by minimum model error''
  paragraphs are redundant, as the same information is already
  displayed in the x-legend for Figures 5 and 6.
}
  { The figures have been redone as subfigures based on Reviewer ones comment. }

\REVIEW{DONE}{In Table 3, what is the unit for the error terms? The
  errors in Figures 7 and 8 are percentages. If errors in Table 3 are
  also percentages please clearly mark. Also, the numbers in Figures 7
  and 8 do not align with numbers in row 1 of Table 3. Please check
  your numbers with care.
}
  {The table has been corrected. See details Reviewer 1}

@@

\REVIEW{DONE}{Please adjust the location of your figures so they do not
  look pages apart from their interpretations.
}
  {The locations of the figures have been adjusted, although LaTeX tries to automatically optimize their
  placement. We have used the !h flag to force placement. If this
  placement is still not appropriate, then please let us know. 
   We have also
  removed images in order to facilitate better placement. The
  document is submitted in draft mode which might also influence the
  placement. We have tried the document in final mode with the LaTeX
  style and it seems that the placement is now fixed as we removed images.}

\REVIEW{DONE}{If I am reading it correctly, Figures 9 and 11 are only
  the “zoom in” versions of partial information from Figure
  13. Similarly for the right column. Why waste space on two more
  figures when you could make Figures 13 and 14 bigger and explain the
  findings?
}
  {As we noted in an earlier comment, now we only include the
  Top 10 instead of Top 1, and Top 5, to ensure -- as per the suggestion of the
  reviewers -- that these images are combined. However we still separated
  the deaths from the cases in the figure to ease the readers' ability to view
  more details. However, we can easily replace the image with the ones
  provided earlier if required. This is a minor change and we welcome any further feedback on this from the reviewers.}

\REVIEW{DONE}{The last paragraph of Section 5.2 suits the
  conclusion/discussion for future work section, not in the middle of
  a paper.
}
  { This paragraph has been moved to the conclusion.}

\REVIEW{DONE}{Section 5.3, the authors mentioned two way of constructing
  covariates. Since the second way was more complicated while yielding
  no better performance the authors did not use it. Then it can be
  removed without damaging the storyline.
}
  { Based on the reviewers suggestion we removed section 5.3 so we could focus on the previous data analysis only. }

\REVIEW{DONE}{The description “We renormalized, so the two basic
  features (following day) contributed 50\% of the loss function” is
  hard to interpret. Please elaborate.
}
    { Based on the reviewers suggestion we removed section 5.3 so we could focus on the previous data analysis only. }

\REVIEW{DONE}{Paragraph 2 on page 14, “Selu, Relu and Tanh activations
  gave very similar results while Sigmoid activation was much
  worse. However we still use a Sigmoid function...” is very
  counter-intuitive. Isn’t the point of trying different activation
  functions to better fit the data?  16. Paragraph 3 on page 14, “The
  basic COVID-19 daily data is extensive but compared two approaches
  with and without renormalization, but in this case, we ...” The
  sentence does not read correct. Please check your sentences with
  care!
}
  { Based on the reviewers suggestion we removed section 5.3 so we could focus on the previous data analysis only. }

\REVIEW{DONE}{How to interpret the “significant positive correlation” in
  Section 6 from the practical perspective? Also, the term
  “evolution function” made its first appearance in Section 6 without
  any definition or explanation; I suggest either removing the whole
  Section 6, or extend it with clearer definition of evolution
  function.  18. The first two paragraphs do not read like to be in a
  “Conclusion” section. Rather, they better fit into the introduction.
}
  { Based on the suggestion of the reviewer we deleted Section 6}

### 2021-01-12 vonLaszewski-covid.tex
Updates from Overleaf

\documentclass[letterpaper, inpress]{jds} % use this for production
%\documentclass[a4paper, review]{jds}      % use this for review
\usepackage{lineno}
\linenumbers

@@

% table managemet

@@

In this study, we used a LSTM algorithm for our predictions
\citep{www-keras-lstm,Hochreiter1997-dk}. An LSTM is a Recurrent
Neural Network (RNN) \citep{Rumelhart1986-li} with feedback
connections allowing the use of data input sequences to predict data
output sequences. LSTMs have been applied to many application areas
from handwriting \citep{Graves2009-qb}, image, feature detection, and
time series prediction \citep{Schmidhuber2005-oy}. LSTM's have an
internal state that is used to prevent the vanishing gradient problem
\citep{Hochreiter1991-mp} during the training of RNNs.

Different variants of LSTM algorithms exist. Ours is based on an LSTM
cell depicted in Figure \ref{fig:lstm}.  It has an input gate, output
gate, and a forget gate. The cell is maintaining how a subsequent
value is calculated. This includes (a) how input values are
influencing the cell via the input gate, (b) how the forget gate
influences a memory value within the cell via the forget gate, and (c)
how the output gate influences the output activation of the LSTM
unit. A cell has a number of inputs and outputs that are
weighted. During a training step, these weights are learned to be
reused in a prediction process between input and output sequences.

\todo{USE NEW REF: \cite{lstm-explained}}
\todo{arrow on tanh is wrong}

\begin{figure}[h!]
\begin{minipage}[b]{0.6\textwidth}
    \centering
    \includegraphics[width=0.6\columnwidth]{images/lstm.pdf}
    \caption{LSTM Unit Diagram}
    \label{fig:lstm}
\end{minipage}
\ \
\begin{minipage}[b]{0.3\textwidth}
\vspace{-3cm}
\begin{equation*}
\begin{split}
    f_t &= \sigma(W_f x_t + U_f h_{t-1} + b_f) \\
    i_t &= \sigma(W_i x_t + U_i h_{t-1} + b_i) \\
    o_t &= \sigma(W_o x_t + U_o h_{t-1} + b_o)\\
    \tilde{c}_t &= tanh(W_c x_t + U_c h_{t-1} + b_c)\\
    c_t &= f_t \circ c_{t-1} +i_t \circ \tilde{c}_t \\ 
    h_t &= o_t \circ tanh(c_t) 
    \label{eq:lstm_equations}
\end{split}
\end{equation*}
\caption{LSTM Equations}
\label{fig:eq}
\end{minipage}
\end{figure}

\newcommand{\Rh}{\mathbb{R}^{h}}

We have summarized the equations used in LSTM in
Figure~\ref{fig:eq}. We denote variables in lower case and matrices in
upper case letters. $W_{q}$ and $U_{q}$ contain the weights of the
input and recurrent connections, where the subscript denotes either
the input gate $i$, the output gate $o$, the forget gate $f$ or the
memory cell $c$.  It is important to note that $c_{t}\in \mathbb
{R}^{h}$ is not just one cell of one LSTM unit, but contains $h$ LSTM
units' cells, representing a vector. To summarize our variables we
use:
$\circ$ is the Hadamard product,
$\sigma$ is the sigmoid function,
$d$ the number of input features,
$h$ the number of hidden units,
$x_{t}\in \mathbb R^d$ the input vector to the LSTM unit,
$f_t\in \Rh$ the forget gate's activation vector,
$i_{t}\in \Rh$ the input/update gate's activation vector,
$o_{t}\in \Rh$ the output gate's activation vector,
$h_{t}\in \Rh$ the hidden state vector (e.g., the output vector of the LSTM unit),
$\tilde{c}_t\in \Rh$ the cell input activation vector,
$c_{t}\in \Rh$ the cell state vector,
$W\in \mathbb{R}^{h\times d}$, $U\in \mathbb{R} ^{h\times h}$, and
$b\in \Rh$ are the weight matrices and bias vector parameters which
need to be learned during training.

@@

In this paper, we adopt this approach to describe the
time dependence of COVID-19 infection and fatality counts in different
regions. 
We intend to
process the two time-series (infections, fatalities/deaths) associated with
each city plus the covariates (fixed in time but dependent on region)
to build a model for the time evolution of the data. There is a
nontrivial architecture design choice as to the method of handling the
time-independent covariates. We could combine a recurrent network for
the time series with some fixed network for the covariates. However,
that does not fit with our concept of deriving an evolution operator
for COVID-19 data that naturally depends on the covariates for each
city. Thus instead, we feed the covariates into the recurrent network.

@@

\begin{requirement}{\bf Expandable API.} \normalfont
  Assuming that new models and other analysis algorithms will be
  developed over time, AICov must allow the integration of such
  APIs. \Solution To address this requirement, we are developing an
  abstract API for data, analysis, and metadata parameter
  adjustments. External services can be integrated with the help of
  REST (REpresentational State Transfer) services. The REST services
  are automatically generated from function specifications.
\end{requirement}

@@

\begin{requirement}{\bf Compute Resource Mapping.} \normalfont
  For other uses of AICov, we need to integrate a flexible resource
  utilization framework. \Solution To address this requirement, we
  leverage from our earlier work and utilize through Cloudmesh and
  National Institute of Standards and Technology (NIST) Big Data
  Working Group (NBDIF) definitions cloud services into the
  architecture \citep{las-19-nist}. This includes containers but also
  infrastructure services in the cloud. Through these service
  integrations, we also provide access to Graphics Processing Units
  (GPUs).
\end{requirement}

\begin{requirement}{\bf Convenient Interfaces.} \normalfont
  As we have a wide variety of users, we need to enable interfaces
  that are used by the various communities. \Solution To address this
  requirement, we provide a number of APIs, REST services and make
  them available via Jupyter Notebooks. In addition, we developed some
  custom widgets for the notebooks that are specifically targeted
  towards data integration, parameter manipulation, and visualization
  of the results
\end{requirement}

@@

Lastly, the total number of general acute care, critical access, and
military hospitals within each county are included in the data. Such
variables included the number of relevant hospitals per county, and
the estimated number of beds (total known bed counts added to the
number of hospitals in the county with missing data times the average
number of beds per hospital in that state) per 1,000 people using the
American Hospital Directory~\citep{www-ahd}.

The list of covariates is presented in Table \ref{tab:risk-factors}.

@@

\begin{question} \normalfont What impact do we see when we introduce many risk factors? See Section \ref{sec:lstm-many}.
\label{q:4}
\end{question}

@@

To establish the feasibility of using deep learning, we have, in our
analysis, verified that through deep learning we can recreate the
predictions identified by a model-based approach using empirical
fits. In this comprehensive analysis, we obtain predictions through
deep learning while using data generated from epidemiological models
as discussed in \citep{marsland20-covid-paper}. The time series we used
was 100 days long, and a multi-layered LSTM
recurrent network was used. Our prediction approach differs by
learning not only from the demographics (fixed data for each city) and
time-dependent data but by integrating the population model for the
underlying prediction, as shown in Figure \ref{fig:magic-1}. Such model
predictive integration capability is important in any application with
multiple time scales. For example, this allows us to integrate
multi-scale time effects into the forecast, which could address the
combination of general forecasts and the next time steps of choice,
such as days, weeks, months, or longer. For this analysis, we used 37
of the 110 cities with reliable empirical fits (that we derived from
an epidemiological model and not deep learning
\citep{marsland20-covid-paper}) for the cases and deaths data up to
April 15, 2020. Our analysis identified a sophisticated single deep
learning time evolution operator that can describe these 37 separate
data sets, and smooth fitted data leads to very accurate deep learning
descriptions via LSTM as depicted in both Figure \ref{fig:magic-1}
that are smaller than 1\% from the model predictions.

@@

\subsection{Real time-Data Covariate LSTM Prediction}
\label{sec:lstm-covariate}

In this section, we establish that using deep learning can not only be
applied to model predictions from real time data. As we want to
evaluate the influence of the risk factors on our prediction framework
using not only models but deep learning algorithms, we have devised a
parameter sweep framework based on LSTM that conducts the prediction
for the 110 cities while also considering the risk factors as listed
in Table \ref{tab:risk-factors}. This framework can query any
combination of risk factors as well as not including any, in which
case it defaults to the traditional LSTM algorithm as introduced in
Section \ref{sec:lstm-theory}. Hence, we will answer Questions
\ref{q:2} and \ref{q:3}.

To answer these questions, we chose for our next analysis a date range
between 2020-02-01 and 2020-05-25. We start with a basic LSTM operator
with two layers of LSTM, and initial and final fully connected layers
\citep{Kadupitiya2020-zq}.  For LSTM, we have selected the activation
function based on Rectified Linear Units (RELU). For the last layer
sigmoid is chosen. A drop out rate of 0.2 is used, we use 16 input
nodes, and have a batch size of 110. The maximum epoch is chosen to be
200. The number of nodes internally is 32. The maximum total number of
samples is 12540 when using 3 days as input length. We use a two-layer
LSTM, as shown in Table \ref{tab:model}. Within our system, we have
the ability to set a number of input days. For this analysis, we have
chosen three different input lengths, namely 3, 4, and 5 days. The
term ``None'' in the output shape means this dimension is
variable. For our model it is the batch size.  \todo{What is the batch
size.}

@@

\begin{figure}[!p]
    \centering
    \includegraphics[width=1.0\textwidth]{images/boxwhisker/boxplot_casesNEW.pdf}
    \vspace{-1cm}
    \caption{Influence of risk factors on the accuracy of the
      prediction of COVID-19 cases. The y-axis represents the
      cumulative error over all input data for the cities.}
    \label{fig:box-cases}
    \bigskip

    \includegraphics[width=1.0\textwidth]{images/boxwhisker/boxplot_deathNEW.pdf}
    \vspace{-1cm}
    \caption{Influence of risk factors on the accuracy of the
      prediction of COVID-19 death. The y-axis represents the
      cumulative error over all input data for the cities.}
    \label{fig:box-death}
\end{figure}

@@

We show the errors and the association with a risk factor in Table
\ref{tab:top-1}, where cum\_error represents the cumulative error of
the prediction. We sorted it by the cumulative error for cases. We see
that in our experiment, almost all risk factors lead to an increased
model prediction accuracy. As a result, we note that by including
factors such as population density, physical health Insurance,
population breakdown by ethnicity, number of hospitals, and diabetes,
it leads to better overall predictions for cases. We also see that
many of these factors perform better on average.

>>>>

%\documentclass[letterpaper, inpress]{jds} % use this for production
\documentclass[a4paper, inpress]{jds} % use this for production
%\documentclass[a4paper, review]{jds}      % use this for review

% spacing
% outcomment this for production
\usepackage{setspace}
\usepackage{lineno}
\linenumbers
\doublespace
% above needs to be commented out for production

@@

% table management

@@

In this study, we base our work on a traditional LSTM algorithm for our predictions
\citep{www-keras-lstm,Hochreiter1997-dk}. An LSTM is a Recurrent
Neural Network (RNN) \citep{Rumelhart1986-li} with feedback
connections allowing the use of data input sequences to predict data
output sequences. LSTMs have been applied to many application areas
from handwriting \citep{Graves2009-qb}, image, feature detection, and
time series prediction \citep{Schmidhuber2005-oy}. LSTM's have an
internal state that is used to prevent the vanishing gradient problem
\citep{Hochreiter1991-mp} during the training of RNNs.

Different variants of LSTM algorithms exist. Ours is based on an LSTM
cell depicted in Figure \ref{fig:lstm}.

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.6\columnwidth]{images/lstm.pdf}
    \caption{LSTM Unit Diagram}
    \label{fig:lstm}
\end{figure}

It has an input gate, output
gate, and a forget gate. The cell is maintaining how a subsequent
value is calculated. This includes (a) how input values are
influencing the cell via the input gate, (b) how the forget gate
influences a memory value within the cell via the forget gate, and (c)
how the output gate influences the output activation of the LSTM
unit. A cell has a number of inputs and outputs that are
weighted. During a training step, these weights are learned to be
reused in a prediction process between input and output sequences.
The LSTM network will have a number of input values represented as  vector x and a number of output values in the last layer of the network predicting a number of future values \citep{lstm-explained}. For our experiments this includes the numbers of days we use as input to our prediction and the number of days we use as output of our prediction for a single cell. In our implementation the number for inputs and outputs can be adjusted. We repeat this prediction over all input sequences to create the output forecast sequences. Next we list the equations (\ref{eq:first}) - (\ref{eq:last}) used in LSTM.

\newcommand{\RH}{\mathbb{R}^{H}}
\newcommand{\RD}{\mathbb{R}^{D}}
\begin{align}
    f_t &= \sigma_g(W_f x_t + U_f h_{t-1} + b_f) \label{eq:first} \\
    i_t &= \sigma_g(W_i x_t + U_i h_{t-1} + b_i) \\
    o_t &= \sigma_g(W_o x_t + U_o h_{t-1} + b_o) \\
    \tilde{c}_t &= \sigma_c(W_c x_t + U_c h_{t-1} + b_c) \\
    c_t &= f_t \circ c_{t-1} +i_t \circ \tilde{c}_t  \\
    h_t &= o_t \circ \sigma_h(c_t) \label{eq:last}
\end{align}

We use in the euqations $\circ$ as the Hadamard product,
$\sigma_g$ is the sigmoid function for the gate,
$D$ the number of days as input features, and
$H$ the number of hidden layers.
Inputs and outputs are defined by 
$x_{t} \in \RD$ the input vector to the LSTM unit, and 
$h_{t}\in \RH$ the hidden state vector e.g., the last of which is the output vector of the LSTM. In our case the output contains the days we predict based on the input features. Other variables include
$f_t \in \RH$ is the forget gate's activation vector, 
$i_t \in \RH$ is the input/update gates activation vector, 
$o_{t}\in \RH$ output gate's activation vector, 
$\tilde{c}_{t}\in \RH$ the cell input activation vector
$c_{t}\in \RH$ the cell state vector,  
$\sigma_c$ is the cell activation function, 
$W\in \mathbb {R} ^{H\times D}$, $U\in \mathbb {R} ^{H\times H}$ and $b\in \RH$ being the weight matrices and bias vector parameters which are learned during training.
In particular we denote $W_{q}$ and $U_{q}$ contain the weights of the
input and recurrent connections, where the subscript denotes either
the input gate $i$, the output gate $o$, the forget gate $f$ or the
memory cell $c$. For a more detailed discussion about LSTM's we refer to \cite{lstm-explained}.

@@

In this paper, we adopt this approach to describe the
time dependence of COVID-19 infection and fatality counts in different
regions. 
We intend to
process the two time-series (infections, fatalities/deaths) associated with
each city plus the covariates (fixed in time but dependent on region)
to build a model for the time evolution of the data. There is a
nontrivial architecture design choice as to the method of handling the
time-independent covariates. We could combine a recurrent network for
the time series with some fixed network for the covariates. However,
that does not fit with our concept of deriving an evolution operator
for COVID-19 data that naturally depends on the covariates for each
city. Thus instead, we feed the covariates into the recurrent network as input in addition to the days.

@@

\begin{requirement}{\bf Expandable API.} \normalfont
  Assuming that new models and other analysis algorithms will be
  developed over time, AICov must allow the integration of such
  models through programs and APIs. \Solution To address this 
  requirement, we are developing an
  abstract API for data, analysis, and metadata parameter
  adjustments. External services can be integrated with the help of
  REST (REpresentational State Transfer) services. The REST services
  are automatically generated from function specifications \citep{las-19-openapi-function}.
\end{requirement}

@@

\begin{requirement}{\bf Compute Resource Mapping.} \normalfont
  For other uses of AICov, we need to integrate a flexible resource
  utilization framework. \Solution To address this requirement, we
  leverage from our earlier work and utilize through Cloudmesh and
  National Institute of Standards and Technology (NIST) Big Data
  Working Group (NBDIF) definitions cloud services into the
  architecture \citep{las-19-nist}. This includes containers but also
  infrastructure services in the cloud. Through the service
  integration, we also provide access to Graphics Processing Units
  (GPUs).
\end{requirement}

\begin{requirement}{\bf Convenient Interfaces.} \normalfont
  As we have a wide variety of users, we need to enable interfaces
  that are used by the various communities. \Solution To address this
  requirement, we provide a number of APIs, REST services and make
  them available via Jupyter Notebooks. In addition, we developed some
  custom widgets for the notebooks that are specifically targeted
  towards data integration, parameter manipulation, and visualization
  of the results.
\end{requirement}

@@

Lastly, the total number of general acute care, critical access, and
military hospitals within each county are included in the data. Such
variables included the number of relevant hospitals per county, and
the estimated number of beds (total known bed counts added to the
number of hospitals in the county with missing data times the average
number of beds per hospital in that state) per 1,000 people using the
American Hospital Directory~\citep{www-ahd}.
The list of covariates is presented in Table \ref{tab:risk-factors}.

@@

\begin{question} \normalfont Are there some impact factors that influence the prediction more than others? See Section \ref{sec:lstm-covariate}.
\label{q:4}
\end{question}

\begin{comment}
\begin{question} \normalfont What impact do we see when we introduce many risk factors? See Section \ref{sec:lstm-many}.
\label{q:4}
\end{question}
\end{comment}

@@

To establish the feasibility of using LSTM, we have, in our
analysis, verified that through deep learning we can recreate the
predictions identified by a model-based approach using empirical
fits. In this comprehensive analysis, we obtain predictions through
deep learning while using data generated from epidemiological models
as discussed in \citep{marsland20-covid-paper}. The time series we used
was 100 days long, and a multi-layered LSTM
recurrent network was used. Our prediction approach differs by
learning not only from the demographics (fixed data for each city) and
time-dependent data but by integrating the population model for the
underlying prediction, as shown in Figure \ref{fig:magic-1}. Such model
predictive integration capability is important in any application with
multiple time scales. For example, this allows us to integrate
multi-scale time effects into the forecast, which could address the
combination of general forecasts and the next time steps of choice,
such as days, weeks, months, or longer. For this analysis, we used 37
of the 110 cities with reliable empirical fits (that we derived from
an epidemiological model and not deep learning
\citep{marsland20-covid-paper}) for the cases and deaths data up to
April 15, 2020. Our analysis identified a sophisticated single deep
learning time evolution operator that can describe these 37 separate
data sets, and smooth fitted data leads to very accurate deep learning
descriptions via LSTM as depicted in Figure \ref{fig:magic-1}
that are smaller than 1\% from the model predictions.

@@

\subsection{Realtime-Data Covariate LSTM Prediction}
\label{sec:lstm-covariate}

In this section, we establish that using deep learning can not only be
applied to model predictions but also from real time data. As we want to
evaluate the influence of the risk factors on our prediction framework
using not only models but deep learning algorithms, we have devised a
parameter sweep framework based on LSTM that conducts the prediction
for 110 cities while also considering the risk factors as listed
in Table \ref{tab:risk-factors}. This framework can query any
combination of risk factors as well as not including any, in which
case it defaults to the traditional LSTM algorithm as introduced in
Section \ref{sec:lstm-theory}. Hence, we will answer Questions
\ref{q:2} and \ref{q:3}.

To answer these questions, we chose for our next analysis a date range
between 2020-02-01 and 2020-05-25. We start with a basic LSTM operator
with two layers of LSTM, and initial and final fully connected layers
\citep{Kadupitiya2020-zq}.  For LSTM, we have selected the activation
function based on Rectified Linear Units (RELU). For the last layer
sigmoid is chosen. A drop out rate of 0.2 is used, we use 16 input
nodes, and have a batch size of 110. The  
batch size is the subset size of the training sample (e.g. 110 out of 12540 for a 3 day input length) that is used to train the LSTM during its learning process. The maximum epoch is chosen to be
200. This is the number of steps after we interrupt the learning process. The number of nodes internally is 32. The maximum total number of
samples is 12540 when using 3 days as input length. We use a two-layer
LSTM, as shown in Table \ref{tab:model}. Within our system, we have
the ability to set a number of input days. For this analysis, we have
chosen three different input lengths, namely 3, 4, and 5 days. The
term ``None'' in the output shape means this dimension is
variable.

@@

\begin{figure}[!p]
    \centering
    \includegraphics[width=1.0\textwidth]{images/boxwhisker/boxplot_casesNEW2.pdf}
    \vspace{-1cm}
    \caption{Influence of risk factors on the accuracy of the
      prediction of COVID-19 cases. The y-axis represents the
      cumulative error over all input data for the cities.}
    \label{fig:box-cases}
    \bigskip

    \includegraphics[width=1.0\textwidth]{images/boxwhisker/boxplot_deathNEW2.pdf}
    \vspace{-1cm}
    \caption{Influence of risk factors on the accuracy of the
      prediction of COVID-19 death. The y-axis represents the
      cumulative error over all input data for the cities.}
    \label{fig:box-death}
\end{figure}

@@

We show the errors and the association with a risk factors in Table
 \ref{tab:top-1-cases} and \ref{tab:top-1-deaths}. In the tables we include the following columns (a) the Risk Factor (b) the root mean square error of all models for this risk factor (c) the minimum cumulative error by risk factor and (d) the number of days that wer used as input and identified to relate to the model with the minimum cumulative error. The table has in the first column the rank this particular model comperes to all other models with minimal risk factor. 
The rank was obtained by sorting the best model for each risk factor. 
We display two tables. One for cases (Table \ref{tab:top-1-cases}) and one for deaths (Table \ref{tab:top-1-deaths}).

We see that in our experiment, almost all risk factors lead to an increased
model prediction accuracy for cases.  We note that by including
factors such as population density, physical health Insurance,
population breakdown by ethnicity, number of hospitals, and diabetes,
lead to better overall predictions for cases. 
We also see that
many of these factors perform better on average.

\begin{table}[!p]
\caption{Best models for cases for each risk factor and their respective
  errors. The model that uses no risk factor is highlighted in grey. We present the Root Mean Square Error (RMSE), the cumulative error and how many days as input were used for this corresponding model.}
\label{tab:top-1-cases}
\bigskip
\centering
%\resizebox{1.0\textwidth}{!}{ 
\begin{tabular}{rrrrrlrl}
\toprule
 \makecell{Place/\\Rank} & 
\makecell{Risk Factor} &  
\makecell{RMSE\\Cases} &  
  \makecell[r]{Minimum\\Cumu-\\lative \\Error for \\Cases by\\Risk Factor} & 
 \makecell[r]{Number of \\Days as \\Input} \\
\midrule

1	&	POP\_DENSITY\_2010	&	0.055443	&	7.94	&	5	 \\
2	&	PHLTH	&	0.054795	&	8.01	&	3	 \\
3	&	BLACK	&	0.055028	&	8.12	&	4	 \\
4	&	HISPANIC	&	0.056061	&	8.22	&	5	 \\
5	&	NHOSP	&	0.054621	&	8.26	&	3	 \\
6	&	INSURANCE	&	0.054557	&	8.34	&	3	 \\
7	&	CHOLSCREEN	&	0.054712	&	8.37	&	3	 \\
8	&	DIABETES	&	0.054586	&	8.4	&	5	 \\
9	&	STROKE	&	0.05416	&	8.45	&	4	 \\
10	&	CHD	&	0.054027	&	8.46	&	5	 \\
11	&	CHECKUP	&	0.054927	&	8.51	&	5	 \\
12	&	NBEDS	&	0.054146	&	8.54	&	5	 \\
13	&	SVI\_OVERALL	&	0.054795	&	8.58	&	5	 \\
14	&	KIDNEY	&	0.0546	&	8.61	&	5	 \\
15	&	CASTHMA	&	0.055089	&	8.63	&	5	 \\
16	&	BPMED	&	0.054685	&	8.74	&	3	 \\
17	&	LPA	&	0.054323	&	8.75	&	5	 \\
18	&	CSMOKING	&	0.053716	&	8.75	&	5	 \\
19	&	ARTHRITIS	&	0.054053	&	8.83	&	3	 \\
20	&	POVERTY	&	0.055029	&	8.84	&	3	 \\
21	&	ESTBEDS	&	0.054545	&	8.86	&	4	 \\
22	&	MHLTH	&	0.054864	&	8.9	&	3	 \\
23	&	COPD	&	0.055589	&	8.93	&	4	 \\
24	&	SENIOR	&	0.055079	&	8.93	&	3	 \\
25	&	CANCER	&	0.0545	&	8.99	&	3	 \\
26	&	NBEDS/1000	&	0.054397	&	9.05	&	3	 \\
27	&	BPHIGH	&	0.054609	&	9.13	&	4	 \\
\rowcolor{black!5} 28	&	NONE	&	0.054868	&	9.26	&	3	 \\
29	&	HIGHCHOL	&	0.056429	&	9.28	&	3	 \\
30	&	BINGE	&	0.054317	&	9.32	&	5	 \\
31	&	OBESITY	&	0.054469	&	9.5	&	3	 \\
32	&	SVI\_MINORITY	&	0.055094	&	9.54	&	4	 \\
\bottomrule
\end{tabular}
%}
\end{table}
